---
title: "Homework 3"
subtitle: "DA 6233"
author: "Samuel Cynamon | yjj265"
date: "2 November 2023"
output: 
  html_notebook:
    theme: cosmo
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
library(tidyverse)
library(plotly)
library(highcharter)
knitr::opts_chunk$set(echo = TRUE)
```

Before you begin, note that, in the header, the output format of this document is `html_notebook`. When you save this file, it automatically creates another file with the same file name but with `.nb.html` extension in the same directory. This is the file you will submit as your homework solution along with the `.Rmd` file. 

<font color = "red"> 
**Warnings**: 

1) Don't delete the `nb.html` file. 
2) Don't `knit` your `.Rmd` file to `html`. If you want to look at the output, just open the `nb.html` in the browser. Alternatively, click on the "Preview" button on top of the document.

If you delete `nb.html` file, you may have to create a new `.Rmd` file and restart from there. If you knit your `.Rmd` file to `html`, you will not be able to retain any of the interactivity in the plots. *This means the TA will not be able to grade you!*
</font>


The objective of this homework is to give you more practice on interactive visualizations using `plotly` and `highcharter`. 

As always, recreate these visualizations exactly. Q1 uses `plotly` while Q2-Q5 use `highcharter`. 

## Q1 (3 points)

Use `mpg` data set from `ggplot2` to create a static visualization and then use `ggplotly()` to create a limited interactive plot.

**Hint**: You will need to supply only `frame`. No `ids` used.

```{r warning=FALSE, fig.width=9}
gg <- ggplot(mpg, 
             aes(x = cty, y = hwy, color=class)) +
  geom_point(aes(frame = class)) +
  scale_x_continuous() +
  labs(x = '',
       y = 'hwy') +
  #scale_color_manual(values = wesanderson::wes_palette("Moonrise3")) +
  theme(legend.position = "none")+
  guides(color=FALSE)

ggplotly(gg)
```


For the next four questions, you will use [`highcharter`](https://jkunst.com/highcharter/). 

## Q2 (3 points)
This example creates a heatmap similar to the one [shown here](https://jkunst.com/highcharter/articles/highcharter.html).

Use `mpg` data and `hchart()` function. We want to create a heatmap of average highway mileage for different `class` and `cyl`. This plot removes all the observations with five cylinders or with `2seater` class. Also note that I am treating `cyl` as a character (string) variable. This is essential for creating this plot.

I am using `hc_theme_538()`. Furthermore, the default color in the heamap is blue, which I changed using `hc_colorAxis()` function that I used in the Week 10 heatmap. 

```{r fig.width=9, fig.height=6}
heatmap_mpg <- mpg %>% 
  select(cyl, class, hwy) %>% 
  filter(!class %in% c('2seater')) %>% 
  filter(!cyl == 5) %>% 
  mutate(
    cyl_str = as.character(cyl)
  ) %>% 
  select(cyl_str, class, hwy) %>% 
  group_by(cyl_str, class) %>%
  summarize(
    hwy_avg = mean(hwy, na.rm=TRUE)
  ) 

heatmap_mpg %>% 
  hchart(type = "heatmap", hcaes(x = class, y = cyl_str, value = round(hwy_avg,2))) %>%
  hc_add_theme(hc_theme_538()) %>% 
  hc_colorAxis(
    stops = color_stops(colors = rev(c("#000004FF", 
                                   "#56106EFF", 
                                   "#BB3754FF", 
                                   "#F98C0AFF", 
                                   "#FCFFA4FF")))
    ) %>% 
  hc_yAxis(title = list(text = "as.character(cyl)")) 


```

## Q3 (3 points)

In the above plot, the tooltip shows confusing information. Below, I modified the tooltip to present more information. The code is not at all complicated and relies on the tooltip code we used in Week 10.

Next, I removed the X axis title and modified Y axis title. 

Finally, I added a title to the plot. Note how I used four different emojies related to cars. It doesn't matter which car emojis you use as long as they are related to automobiles.


```{r fig.width=9, fig.height=6}

heatmap_mpg %>% 
  hchart(type = "heatmap", hcaes(x = class, y = cyl_str, value = round(hwy_avg,2)), name='Highway Mileage') %>%
  hc_add_theme(hc_theme_538()) %>% 
  hc_colorAxis(
    stops = color_stops(colors = rev(c("#000004FF", 
                                   "#56106EFF", 
                                   "#BB3754FF", 
                                   "#F98C0AFF", 
                                   "#FCFFA4FF")))
    ) %>% 
  hc_title(text = 'Highway Mileage Decreases across all the \U1F697 \U1F699 \U1F3CE \U1F690 as the Number of Cylinders Increases', useHTML = TRUE) %>% 
  hc_yAxis(title = list(text = "Number of Cylinders")) %>% 
  hc_xAxis(title = list(text = "")) %>% 
  hc_tooltip(
    pointFormat = "For class {point.class} with {point.cyl_str} cylinders: {point.value} mpg") 
```


## Q4 (3 points)

For this example, use a randomly selected subset of `diamonds` data set from `ggplot2`:

```{r echo=TRUE}
set.seed(2020)
d1 = diamonds[sample(nrow(diamonds), 1000),]
```

Next use `d1` to create the following plot. 

I have used `hc_theme_flat()` for this plot. **Please use this theme for your plot too!**
You can add a theme to the plot using `hc_add_theme()` function. Wherever the word diamond appeared in the plot, I replaced it with the diamond emoji.

Point colors in this graph are mapped to `clarity`. Check out all the variables in this data set by typing `?diamonds` in the console.

```{r fig.width=9, fig.height=6}
d1 %>% 
  hchart("scatter", hcaes(x = carat, y = price, group=clarity)) %>% 
  hc_add_theme(hc_theme_flat()) %>% 
  hc_title(text = 'Variation in Prices for \U1F48E Increases with Carats', useHTML = TRUE) %>% 
  hc_yAxis(title = list(text = "Price of \U1F48E"), tickInterval = 2500) %>% 
  hc_xAxis(title = list(text = "Weight of \U1F48E in Carats")) 

```


## Q5 (3 points)

Recreate the plot in Q2 using `hchart()`. I used `hc_theme_economist()`. You can use any theme you want. You can check out the themes [here](https://jkunst.com/highcharter/articles/themes.html). I used `economics` dataset from `ggplot2`. Learn more about the variables in the dataset by typing `?economics` in the console.

```{r fig.width=9, fig.height=6}
economics %>% 
  hchart('line', hcaes(x = date, y = unemploy), name="Unemployment") %>% 
  hc_add_theme(hc_theme_economist()) %>% 
  hc_title(text = 'Unemployment Peaked after the Financial Crisis') %>% 
  hc_yAxis(title = list(text = "Unemployment in '000"), tickInterval = 2000) %>% 
  hc_xAxis(title = list(text = "Date"), labels = list(step = 2)) 
```


## Bonus plot (Not graded)

This is the same plot as above except if you hover mouse pointer over the peak of unemployment, the tooltip will show more information. Once again, this is a simple trick and doesn't require any advanced coding. 


```{r fig.width=9, fig.height=6}
economics$Custom = ""
economics["Custom"][economics['unemploy'] == max(economics$unemploy)] <- "Mortgage Crisis \U1F3E6"


economics %>% 
  hchart('line', hcaes(x = date, y = unemploy), name="Unemployment") %>% 
  hc_add_theme(hc_theme_economist()) %>% 
  hc_title(text = 'Unemployment Peaked after the Financial Crisis') %>% 
  hc_yAxis(title = list(text = "Unemployment in '000"), tickInterval = 2000) %>% 
  hc_xAxis(title = list(text = "Date"), labels = list(step = 2)) %>% 
  hc_tooltip(pointFormat="<span style='color:#6794a7'> \U25CF </span>Unemployment: <b>{point.y}</b> <br> <span style = 'color:#ff0000'>{point.Custom}</span>"
            ,useHTML=TRUE)
```


